Coasting on Duality

### 2024-07-10 main.tex
ncsu presentation
Updates from Overleaf

% Main title of the paper
\title [mode = title]{The Correspondence Analysis of Two-Mode Networks: Centrality and Positional Interpretations}                      
% Title footnote mark
% eg: \tnotemark[1]
%\tnotemark[1,2]

@@

% First author
%
% Options: Use if required
% eg: \author[1,3]{Author Name}[type=editor,
%       style=chinese,
%       auid=000,
%       bioid=1,
%       prefix=Sir,
       %orcid=0000-0002-5405-3007,
%       facebook=<facebook id>,
%       twitter=<twitter id>,
%       linkedin=<linkedin id>,
%       gplus=<gplus id>]
\author[1]{Omar Lizardo}[orcid=0000-0001-7511-2910]

@@

% For a title note without a number/mark
%\nonumnote{This note has no numbers. In this work we demonstrate $a_b$
%  the formation Y\_1 of a new type of polariton on the interface
 % between a cuprous oxide slab and a polystyrene micro-sphere placed
 % on the slab.
 % }
% Here goes the abstract
\begin{abstract}
Lorem...
\end{abstract}

@@

%Research highlights
\begin{highlights}
    \item 
    \item 
    \item 
\end{highlights}

% Keywords
% Each keyword is separated by \sep
\begin{keywords}
Generalized similarity \sep Duality \sep Two-mode networks \sep Correspondence Analysis
\end{keywords}

\maketitle
\newpage
\section{Introduction}
While not particularly common, using Correspondence Analysis---hereafter CA---in analyzing two-mode network data has had a somewhat rocky career in the social networks literature. Initially criticized by \citet{borgatti1997network} as a relatively limited and perhaps even inapplicable tool, CA has found various proponents who see it as an important component of the SNA arsenal for two-mode network data analysis, particularly regarding its ability to economically provide synoptic (e.g., ``joint'') graphical representations of structural connectivity patterns across the two-modes \citep{roberts2000correspondence, breiger2000tool, faust2005using}, with the primary aim being to use CA---or its variants like Multiple Correspondence Analysis (MCA)---to ``visually explore'' such networks \citep{d2014use}. This paper contributes to the stream of previous work applying CA to analyze two-mode data. Nevertheless, it departs from the already-mentioned previous efforts in emphasizing aspects of CA for two-mode data analysis in SNA that have not been as emphasized or treated in detail before, moving beyond the focus on data visualization.

Particularly, I show that CA can be thought of as computing a kind of \textit{dual centrality} score on both modes---one related to but not mathematically equivalent to the Bonacich \citeyearpar{bonacich1991simultaneous} two-mode centrality----an approach that, as noted by \citet{van2021correspondence}, has been recently re-invented in some corners of network science under the guise of the ``economic complexity index'' \citep{hidalgo2009building, mealy2019interpreting}. With \citet{van2021correspondence}, I argue that this is a rather restricted interpretation of a more general ``dual centrality'' score that connects to the use of CA in some disciplines (e.g., ecology) as a method of \textit{ordination}---e.g., the discovery of latent one-dimensional orderings among a set of entities. In social network analysis, this applies to conceptualizing the venerable connection between people and the groups they join \citep{breiger1974duality}.

Additionally, I argue that CA can also be used to extract latent \textit{positional} information on people and groups. First, I show that there is a suggestive similarity between the scores obtained from the first dimension of the CA of the two-mode affiliation matrix (the aforementioned dual centrality score) and recent work on ``generalized relational similarity'' in two-mode networks \citep{kovacs2010generalized, lizardo2024two}. As such, CA recovers clusters of entities (e.g., people) linked by their similar connectivity patterns to similar entities in the other mode (e.g., groups). Second, drawing on recent work by \citet{van2021correspondence}, I show that the first \textit{two} CA dimensions provide information on the degree-weighted similarity of nodes in the same mode concerning their connections to nodes in the other mode. Thus, CA can be used as an approach to indirect ``community'' identification in two-mode networks that connects to the use of CA in some disciplines (e.g., computer and information science) as a \textit{clustering} method \citep{zha2001bipartite}---e.g., the discovery of sociometrically similar subsets of entities in affiliation networks.

\subsection{Organization of the Paper}
The rest of the paper is organized as follows. In the next section, I motivate the use of CA as producing a type of Bonacich-style dual centrality via weighted iteration across the two modes of the affiliation matrix. This leads naturally to an abbreviated way of computing the same scores via eigenvector decomposition of the (inverse degree-weighted) one-mode projection of the affiliation matrix for each set of nodes. This approach is computationally and mathematically related but \textit{not} equivalent to the matrix factorization of the one-mode projection matrices, which results in the usual Bonacich dual centrality scores. I then draw on recent work by \citet{van2021correspondence} to show the links between the ordering of nodes along the first CA dimension and Kovacs's generalized relational similarity, and the distance of nodes in the space formed by the first two dimensions and the connectivity-similarity between nodes across modes. I contrast this ordering to that provided by the Bonacich dual centrality scores, which in contrast to the CA ordering---which reveals a dual community partition--- recovers instead the core-periphery structure of the two-mode network. In closing, I provide some pointers on how CA can be used as a method for structural (e.g., centrality and positional) analysis, independently of the usual emphasis on joint graphical displays and visual exploration.

\section{Reflective Centralities in Two-Mode Networks}
In a highly cited piece, \citet{hidalgo2009building} motivated what they saw at the time as a novel way of ranking nodes in a two-mode network---effectively computing a version of two-mode centrality---using what they called at the time a ``reflective'' approach. Hidalgo and Hausmann's original empirical application was to the two-mode country-by-product networks, hence the original baptizing of their approach as one geared to extracting the ``economic complexity'' of countries in the world system (and dually the complexity of given products). Subsequent shows that there is no logical connection between the formal method and this particular application. As such, I introduce it here using the more intuitive---and classical from a social network analysis perspective---case of the duality of persons and groups \citep{breiger1974duality}.

@@

That is, the first-order centrality of people is the row sum of the entries in the affiliation matrix $A$, and the column sums of the same matrix give the first-order centrality of groups. In what follows, I will generally refer to $C^R_p(1)$ as (individual) ``activity'' and $C^R_p(1)$ as (group) ``size.''

As noted by \citet{hidalgo2009building}, the key to the reflective approach is the observation that, once we have these first-order quantities, it is possible to compute ``second-order centralities'' $C^R(2)$ for both people and groups using the (averaged) first-order centralities of the entities in the other mode they are connected to.

@@

\subsection{Empirical Example}
Figure~\ref{fig:pg-reflections} shows the trajectory of the HH reflective centralities for persons (top panel) and groups (bottom panel) in the Southern Women data \citep{davis1941}, shown in Table~\ref{tab:southern}. The rank order of people and groups in the $q^{th}$ centrality is plotted in the y-axis, and the centrality iteration $q$ is plotted on the x-axis.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{Plots/pg-reflections.jpg}
    \caption{Reflective centrality trajectories of persons and groups (even reflections).}
    \label{fig:pg-reflections}
\end{figure}
\section{Correspondence Analysis, Dual Centrality, and Communities in Two-Mode Networks}
The reader may ask what the point of going through all these reflective centralities, since \citet{bonacich1991simultaneous} already developed a dual conception of centrality in two-mode networks based on a very similar idea: Define the centralities of entities in one mode based on the centralities of entities in the other mode to which they are connected. In that paper, Bonacich also considered CA in passing but dismissed its application to centrality rankings. Instead, Bonacich noted that, in analyzing the Southern Women data shown in Table~\ref{tab:southern}, ``[r]ather than centrality, what they [CA scores] seem to capture is membership in the two cliques that attended two different sets of events'' \citeyearpar[164]{bonacich1991simultaneous}.

@@

To see this, recall that the solutions to the following eigensystem give the Bonacich $(C^B)$ two-mode centralities:

@@

Where $A$ is the original affiliation matrix. The unknown $C^B_g$ and $C^B_p$ scores can be obtained from the row and column eigenvectors corresponding to the largest eigenvalue $\lambda$, as given by the singular value decomposition (SVD) of the rectangular affiliation matrix. Importantly, as Faust \citeyearpar[170]{faust1997centrality} notes, it is possible to express Equations~\ref{eq:bon1} and~\ref{eq:bon2} in terms of person and group-specific centralities.

@@

Where $\lambda$ is the (first) eigenvalue corresponding to the eigenvector containing the $C^B_p$ scores. These equations make clear that in the Bonacich two-mode centrality of people is proportional to centralities of the groups they join, and the centrality of groups is proportional to the centralities of their members. These capture the duality property since the centralities of nodes in each mode are given by aggregating their connections to nodes in the other mode. Note the formal similarity between equations~\ref{eq:faust1} and~\ref{eq:faust2} and equations~\ref{eq:Rq_p} and~\ref{eq:Rq_g}. The main difference is that in the reflective equations, a person's centrality is proportional to the \textit{activity-weighted} centralities of the groups they join, and a group's centrality is proportional to the \textit{group size weighted} centrality of the people who are members.

As noted by \citet{mealy2019interpreting} and \citet{van2021correspondence}, Hidalgo and Hausmann's \citeyearpar{hidalgo2009building} reflective centralities can also be obtained as the solution to an eigenvalue (matrix factorization) problem. To see this, recall that, as Bonacich \citeyearpar[157]{bonacich1991simultaneous} notes, we can solve for $C^B_g$ in~\ref{eq:bon2} and $C^B_p$ in~\ref{eq:bon1} and substitute the respective solutions into \ref{eq:bon1} and~\ref{eq:bon2}. Matrix-algebraic reduction of the resulting equations would show that the Bonacich dual centralities can also be obtained as solutions to the eigensystem:

@@

Note the formal similarity (and key differences) between these equations and the Bonacich two-mode centralities in equations~\ref{eq:bon3} and~\ref{eq:bon4}. Both extract individual and group centralities as eigenvectors of the one-mode projection of the original affiliation matrix: $AA^T$ for people and $A^TA$ for groups. The difference is that the reflective centralities pre-multiply the affiliation matrix and its transpose by the inverse of the first-order centralities of the nodes in each mode before factorizing, essentially normalizing the one-mode projection by the degrees of both sets of nodes.

@@

Now, as has been noted by two-mode data analysts in other contexts \citep[e.g.,][]{faust2005using} it turns out that this ``double-pre-weighting'' of each cell entry by the degrees of each mode (e.g., the row and column sums of the original affiliation matrix) is precisely that used for the CA of a two-mode matrix.\footnote{More accurately, cells are weighted by inverse of the square root of the product of the row and column sums \citep[e.g.,][124]{faust2005using}.} And indeed, the limit reflective centralities obtained from equations~\ref{eq:dam1} and~\ref{eq:dam2}---which will be given by the eigenvectors corresponding to the \textit{second} largest eigenvalue of the resulting solution---will be equivalent to the row and column scores corresponding to the first non-trivial dimension (for people and groups respectively) obtained from a simple CA of the original affiliation matrix \citep[398, eq. 9.17]{fouss2016algorithms}.\footnote{As \citet[126]{faust2005using}, the first non-trivial CA dimension is also given by the eigenvector corresponding to the second-largest eigenvalue of $D_p^{-1/2}AD_g^{-1/2}$.} In fact, the method of ``reflections'' is just a re-discovery of the older idea of reciprocal averaging \citep{hill1973reciprocal}, which is yet another way of computing row and column scores for elements in a two-mode matrix that are substantively equivalent to those obtained by (the first non-trivial dimension of) CA \citep{mealy2019interpreting}.

Degree-pre-weighting does not radically alter the nature of the data. Indeed, it corresponds to moving from sums to averages, thus ``adjusting'' for the influence of person-activity and group size---a seemingly perenial issue in two-mode data analysis \citep[159ff]{bonacich1991simultaneous}. This can be seen by the fact that Equations~\ref{eq:mealy1} and~\ref{eq:mealy2} show that, substantively, what the degree pre-weighting does is that, for people, co-memberships count for more in determining interpersonal similarity when the group in question is small than when it is large. On the group side, shared members who do not have many affiliations count more in determining intergroup similarity than those with many affiliations.

Accordingly, it would make no sense to call the Bonacich scores $C^B_p$ and $C^B_g$ from equations~\ref{eq:bon3} and~\ref{eq:bon4} ``centrality scores'' but fail to hold that designation from $C^R_p$ and $C^R_g$, given that the mathematics are not only ``similar'' \citep[162]{bonacich1991simultaneous} but formally identical. In the Bonacich case, a factorization of the one-mode projection of the two-mode network; in the reflective case, factorization of the degree-weighted one-mode projection of the same network. Accordingly, the CA of a two-mode network will return---along the first dimension---a rank-ordered score for nodes in each mode that meets all the conditions for qualifying as a centrality measure for two-mode networks. Particularly, the CA centrality retains the duality property \citep[128]{faust2005using}, as the (average) centrality of people is a function of the (average) centrality of the groups they belong to. The (average) centrality of groups, in turn, is a function of their members' (average) centrality. The person and group score ranks given by the first CA dimension are thus better thought of as \textit{activity and group-size normalized} versions of the familiar Bonacich \citeyearpar{bonacich1991simultaneous} dual centralities for two-mode networks.

@@

\section{Illustrative Analysis of the Southern Women Data}
\subsection{Row and Column Re-ordered Affiliation Matrices}
Figures~\ref{fig:ca-v-bon}(a) and~\ref{fig:ca-v-bon}(b) illustrate the key differences between the two versions of dual centrality $C^R$ and $C^B$. Each panel shows the original Southern Women affiliation matrix but with rows and columns re-ordered according to the first CA dimension in (a) and the \citet{bonacich1991simultaneous} centrality in (b). In the plot, a cell entry is colored red if it has a one in the original affiliation matrix and is white when the corresponding entry is zero. As we can see, the two row-column-reshuffled matrices have appreciably distinct structures, with the $C^R$ re-ordered affiliation matrix having a block-diagonal structure and the $C^B$ re-ordered affiliation matrix having a triangular structure. Accordingly, the $C^R$ centrality rankings reveal a dual community partition between groups of women who selectively attend two groups of events (on the top-right and bottom-left of the plot). The traditional eigenvector centrality re-ordering, on the other hand, reveals a classic core-periphery partitioning \citep{borgatti2000models}, with a group of highly active women (on the bottom-right) who selectively co-participate in highly attended events (on the lower right-hand side of the plot) and less active women (in the upper half) who go to less well-attended events (on the right side). The two centralities extract qualitatively different information from the two-mode network, with $C^R$ geared toward community partitioning and $C^B$ more focused on finding a core of well-connected people and groups.

>>>>

% Main title of the paper
\title [mode = title]{The Correspondence Analysis of Two-Mode Networks Revisited}                      
% Title footnote mark
% eg: \tnotemark[1]
%\tnotemark[1,2]

@@

% First author
%
% Options: Use if required
% eg: \author[1,3]{Author Name}[type=editor,
%       style=chinese,
%       auid=000,
%       bioid=1,
%       prefix=Sir,
       %orcid=0000-0002-5405-3007,
%       facebook=<facebook id>,
%       twitter=<twitter id>,
%       linkedin=<linkedin id>,
%       gplus=<gplus id>]
\author[1]{Omar Lizardo}[orcid=0000-0002-5405-3007]

@@

% For a title note without a number/mark
%\nonumnote{This note has no numbers. In this work we demonstrate $a_b$
%  the formation Y\_1 of a new type of polariton on the interface
 % between a cuprous oxide slab and a polystyrene micro-sphere placed
 % on the slab.
 % }
% Here goes the abstract
\begin{abstract}
This paper reconsiders the use of Correspondence Analysis (CA) in analyzing two-mode network data, highlighting aspects that have not been previously emphasized, going beyond the use of CA as a visualization tool. It argues that CA can be used to compute a dual centrality score on both modes, related to but not mathematically equivalent to the \citet{bonacich1991simultaneous} two-mode centrality. This ``reflective'' centrality score connects the use of CA in two-mode network analysis to its use in other disciplines as a method of ordination. Additionally, I show CA can extract latent positional information on people and groups, with similarities to recent work on ``generalized relational similarity" in two-mode networks. Thus, CA can be used for indirect community or subgroup identification in two-mode networks, connecting to its use in some disciplines as a clustering method. The paper demonstrates these applications of CA, comparing them with the Bonacich dual centrality scores, and provides guidance on using CA for structural analysis in network analysis.
\end{abstract}

@@

%Research highlights
\begin{highlights}
  \item This paper revisits the use of Correspondence Analysis (CA) in analyzing two-mode network data, highlighting its potential beyond data visualization.
  \item CA can compute a dual centrality score on both modes of a two-mode network, related to but not mathematically equivalent to the Bonacich (1991) two-mode centrality. 
  \item CA can extract latent positional information on people and groups, with similarities to recent work on generalized relational similarity" in two-mode networks. 
\end{highlights}

% Keywords
% Each keyword is separated by \sep
\begin{keywords}
Correspondence Analysis \sep Generalized Relational Similarity \sep Duality \sep Two-mode networks \sep Centrality 
\end{keywords}

\maketitle
\newpage
\section{Introduction} \label{sec:intro}
While not particularly common, using Correspondence Analysis---hereafter CA---in analyzing two-mode network data has had a somewhat rocky career in the social networks literature. Initially criticized by \citet{borgatti1997network} as a relatively limited and perhaps even inapplicable tool, CA has found various proponents who see it as an important component of the SNA arsenal for two-mode network data analysis, particularly regarding its ability to economically provide synoptic (e.g., ``joint'') graphical representations of structural connectivity patterns across the two-modes \citep{roberts2000correspondence, breiger2000tool, faust2005using}, with the primary aim being to use CA---or its variants like Multiple Correspondence Analysis (MCA)---to ``visually explore'' such networks \citep{d2014use}. This paper contributes to the stream of previous work applying CA to analyze two-mode data. Nevertheless, it departs from the already-mentioned previous efforts in highlighting aspects of CA for two-mode data analysis in SNA that have not been emphasized or treated in detail before, moving beyond the focus on data visualization.

Particularly, I show that CA can be thought of as computing a kind of \textit{dual centrality} score on both modes---one related to but not mathematically equivalent to the Bonacich \citeyearpar{bonacich1991simultaneous} two-mode centrality----an approach that, as noted by \citet{van2021correspondence}, has been recently re-invented in some corners of network science under the guise of the ``economic complexity index'' \citep{hidalgo2009building, mealy2019interpreting}. With \citet{van2021correspondence}, I argue that this is a rather restricted interpretation of a more general centrality score connecting the use of CA in two-mode network analysis to way CA features in some disciplines also concerned with two-mode data (e.g., ecology) as a method of \textit{ordination}---e.g., the discovery of latent one-dimensional orderings among a set of entities. In social network analysis, this applies to conceptualizing the venerable duality between people and the groups they join \citep{breiger1974duality}.

In addition to its capacity to reveal latent ordinal structure in two-mode data, I also argue that CA can also be used to extract latent \textit{positional} information on people and groups. Particularly, there is a suggestive similarity between the scores obtained from the first dimension of the CA of the two-mode affiliation matrix (the aforementioned dual centrality score) and recent work on ``generalized relational similarity'' in two-mode networks \citep{kovacs2010generalized, lizardo2024two}. As such, CA recovers clusters of entities (e.g., people) linked by their similar connectivity patterns to similar entities in the other mode (e.g., groups). Thus, CA can be used as an approach to indirect ``community'' identification in two-mode networks connecting to how CA is used in some disciplines (e.g., computer and information science) as a \textit{clustering} method \citep{zha2001bipartite}, namely, the discovery of sociometrically similar subsets of entities in affiliation networks. In this way, CA emerges as a tool central to the usual network-analytic tasks and not just as a visualization or data-summarization tool.

\subsection{Organization of the Paper} \label{subsec:org}
The rest of the paper is organized as follows. In the next section, I motivate the use of CA as producing a type of Bonacich-style dual centrality via weighted iteration across the two modes of the affiliation matrix. This leads naturally to an abbreviated way of computing the same scores via eigenvector decomposition of the (inverse degree-weighted) one-mode projection of the affiliation matrix for each set of nodes. This approach is computationally and mathematically related but \textit{not} equivalent to the eigenvalue decomposition of the one-mode projection matrices, which results in the usual Bonacich dual centrality scores. I then draw on recent work by \citet{van2021correspondence} to show the links between the ordering of nodes along the first CA dimension and Kovacs's generalized relational similarity, and the distance of nodes in the space formed by the first two dimensions and the connectivity-similarity between nodes across modes. I contrast this ordering to that provided by the Bonacich dual centrality scores, which in contrast to the CA ordering---which reveals a dual community partition--- recovers the core-periphery structure of the two-mode network instead. In closing, I provide some pointers on how CA can be used as a method for structural (e.g., centrality and positional) analysis, independently of the usual emphasis on joint graphical displays and visual exploration.

\section{Reflective Centralities in Two-Mode Networks} \label{sec:ref2mode}
In a highly cited piece, \citet{hidalgo2009building} motivated what they saw at the time as a novel way of ranking nodes in a two-mode network---effectively computing a version of two-mode centrality---using what they called at the time a ``reflective'' approach. Hidalgo and Hausmann's original empirical application was to the two-mode country-by-product networks, hence the original baptizing of their approach as one geared to extracting the ``economic complexity'' of countries in the world system (and dually the complexity of given products). Subsequent work shows that there is no logical connection between the formal method and this particular application since the approach proposed can be used to analyze any two-mode data matrix. As such, I introduce it here using the more intuitive---and classical from a social network analysis perspective---case of the duality of persons and groups \citep{breiger1974duality}.

@@

That is, the first-order centrality of people is the row sum of the entries in the affiliation matrix $A$, and the column sums of the same matrix give the first-order centrality of groups. As noted by \citet{hidalgo2009building}, the key to the reflective approach is the observation that, once we have these first-order quantities, it is possible to compute ``second-order centralities'' $C^R(2)$ for both people and groups using the (averaged) first-order centralities of the entities in the other mode they are connected to.

@@

\subsection{Empirical Example} \label{subsec:ex}
Figure~\ref{fig:pg-reflections} shows the trajectory of the HH reflective centralities for persons (top panel) and groups (bottom panel) in the Southern Women data \citep{davis1941}, shown in Table~\ref{tab:southern}. The rank order of people and groups in the $q^{th}$ centrality is plotted on the y-axis, and the centrality iteration $q$ is plotted on the x-axis.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{Plots/pg-reflections.jpg}
    \caption{Reflective centrality trajectories of persons and groups (even reflections).}
    \label{fig:pg-reflections}
\end{figure}

\section{Correspondence Analysis, Dual Centrality, and Communities in Two-Mode Networks} \label{sec:ca}
The reader may ask what the point of going through all these reflective centralities, since \citet{bonacich1991simultaneous} already developed a dual conception of centrality in two-mode networks based on a very similar idea: Define the centralities of entities in one mode based on the centralities of entities in the other mode to which they are connected. In that paper, Bonacich also considered CA in passing but dismissed its application to centrality rankings. Instead, Bonacich noted that, in analyzing the Southern Women data shown in Table~\ref{tab:southern}, ``[r]ather than centrality, what they [CA scores] seem to capture is membership in the two cliques that attended two different sets of events'' \citeyearpar[164]{bonacich1991simultaneous}.

@@

To see this, recall that the solutions to the following system of linear equations give the Bonacich $(C^B)$ two-mode centralities:

@@

Where $A$ is the original affiliation matrix. Equations~\ref{eq:bon1} and~\ref{eq:bon2} have the typical form of a linear eigensystem, which means the unknown $C^B_g$ and $C^B_p$ scores can be obtained from the row and column eigenvectors corresponding to the largest eigenvalue $\lambda$, obtained from the singular value decomposition (SVD) of the rectangular affiliation matrix $A$. Importantly, as Faust \citeyearpar[170]{faust1997centrality} notes, it is possible to express Equations~\ref{eq:bon1} and~\ref{eq:bon2} in terms of person and group-specific centralities.

@@

Where $\lambda$ is the (first) eigenvalue corresponding to the eigenvector containing the $C^B_p$ scores. These equations make clear that in the Bonacich two-mode, the centrality of people is proportional to the centralities of the groups they join, and the centrality of groups is proportional to the centralities of their members. These capture the duality property since the centralities of nodes in each mode are given by aggregating their connections to nodes in the other mode. Note the formal similarity between equations~\ref{eq:faust1} and~\ref{eq:faust2} and equations~\ref{eq:Rq_p} and~\ref{eq:Rq_g}. The main difference is that in the reflective equations, a person's centrality is proportional to the \textit{activity-weighted} centralities of the groups they join, and a group's centrality is proportional to the \textit{group size weighted} centrality of the people who are members. We will return to this crucial point later.

\subsection{The Bonacich Two-Mode Centrality as a Reflective Centrality} \label{subsec:bonref}
To see the connection between the reflective and Bonacich two-mode centralities more clearly, we can motivate the Bonacich two-mode centrality using the same ``reflective'' approach we used to introduce the HH reflective centrality in Section~\ref{sec:ref2mode}. Admittedly, this is a somewhat unorthodox way of presenting the eigenvector centrality for two-mode networks~\citep{bonacich1991simultaneous}, but it will help clarify the similarities and differences between the HH and the Bonacich approaches. Accordingly, starting with Equations~\ref{eq:R1_p} and~\ref{eq:R1_g}, we can define a second-order ``Bonacich-reflection'' on both the persons and groups using the formulas:

\begin{equation}
   C^B_p(2) = \sum_g a_{pg}C^R_g(1)
   \label{eq:BR2_p}
\end{equation}

\begin{equation}
   C^B_g(2) = \sum_p a_{pg}C^R_p(1)
   \label{eq:BR2_g}
\end{equation}

Equation~\ref{eq:BR2_p} says that people are central when the sum of the number of members of the groups they belong to is large. Equation~\ref{eq:BR2_g} says groups are central when the sum of the number of memberships of the people who belong to them is large.

As before, we can keep on going and define a third-order reflection using the formulas:

\begin{equation}
   C^B_p(3) = \sum_g a_{pg}C^R_g(2)
   \label{eq:BR3_p}
\end{equation}

\begin{equation}
   C^B_g(3) = \sum_p a_{pg}C^R_p(2)
   \label{eq:BR3_g}
\end{equation}

Equation~\ref{eq:BR3_p} says that people are central when the sum of the sum of the number of memberships held by the people who belong to the groups they belong to is large. Equation~\ref{eq:BR3_g} says that groups are central when the sum of the sum of the number of members of the groups their members belong to is large. Once again, we can keep going and define fourth order, fifth order, and higher reflections $C^R_p(4), C^R_p(4) \ldots C^R_p(q)$, where the centralities of nodes in one mode are based on the sums, of the sums, of the sums, of the centralities of nodes in the other mode.

More generally, and in parallel with equations~\ref{eq:Rq_p} and~\ref{eq:Rq_g}, the reflective Bonacich centralities for persons and groups are given by:

\begin{equation}
   C^B_p(q) = \sum_g a_{pg}C^R_g(q-1)
   \label{eq:BRq_p}
\end{equation}

\begin{equation}
   C^B_g(q) = \sum_p a_{pg}C^R_p(q-1)
   \label{eq:BRq_g}
\end{equation}

This just says that the Bonacich reflective centrality at step $q$ is just the sum of the group centralities at step $q-1$ (for people) and the sum of the person centralities at step $q-1$ (for groups).

Of course, it is evident that these sum of sums would diverge to a bigger and bigger quantity at each step $q$. To prevent this and guarantee convergence, we normalize the vector of reflective Bonacich centralities for persons and groups at each step $q > 1$ before calculating the subsequent sum at step $q+1$ as follows:

\begin{equation}
   C^B_p(q) = \frac{C^B_p(q)}{||C^B_p(q)||_2}
   \label{eq:BRq_pn}
\end{equation}

\begin{equation}
   C^B_g(q) = \frac{C^B_g(q)}{||C^B_g(q)||_2}
   \label{eq:BRq_gn}
\end{equation}

Where the denominator in~\ref{eq:BRq_pn} and~\ref{eq:BRq_gn} is the Euclidean vector norm.\footnote{For any vector $\mathbf{x}$ of length $n$, the $L_2$ norm is given by: $||\mathbf{x}||_2 = \sqrt{\sum_i^n x_i^2}$.} The normalization will prevent divergence of the sum of centralities for persons and groups, formalizing the weaker assumption of \textit{proportionality} between the centrality of each set of nodes and the sum of the centralities of the nodes in the other mode to which they are connected rather than the stronger assumption of strict equivalence \citep{bonacich_lloyd01}. Furthermore, the normalization guarantees convergence and the ``freezing'' of the sums around steady values after a few iterations. These values will be equivalent (up to rounding error) to the (absolute value) of the dominant row and column eigenvectors of $\mathbf{A}$ as given in~\ref{eq:bon1} and~\ref{eq:bon2}.\footnote{Note that dividing by any vector norm---e.g., the $L_1$ or max norm---will prevent divergence and return scores perfectly correlated to the Bonacich eigenvector approach. Dividing by the $L_2$ norm returns scores that are \textit{exactly} the same, save for rounding error, as the absolute value of the dominant left and right eigenvectors of the affiliation matrix.} In fact, the iterative (normalized sum of sums) approach is one way of computing the leading eigenvectors of a rectangular matrix (e.g., the ``power'' method of \citet{mises1929praktische}).

This exercise establishes that there is more than a superficial similarity between the HH ``reflective'' centralities and the Bonacich two-mode centralities. In fact, both can be seen as instantiating an underlying reflective model of how centrality is distributed in the two-mode network, with the Bonacich approach presuming that centrality points are distributed equally across nodes in each mode regardless of their own centrality (both large and small degree nodes distribute the same ``amount'' of centrality to others) and the HH reflective approach normalizing by the centrality of nodes in each mode, so that nodes distribute a given centrality quantum that is proportional to their first-order degree centrality, with large degree nodes having less centrality to distribute than low degree nodes.

\subsection{HH Reflective Centrality as an Eigenvector Centrality} \label{subsec:refeigen}
Since iterating through normalized sums of sums is one way of obtaining the Bonacich two-mode centralities, it would be surprising if the equilibrium values of the HH reflective iterations were not themselves the solution to some eigenvalue decomposition problem. As has been noted recently by \citet{mealy2019interpreting} and \citet{van2021correspondence}, the Hidalgo and Hausmann's \citeyearpar{hidalgo2009building} reflective centralities can indeed be obtained directly (without iterations) as the solution to an eigenvalue decomposition problem.

To see this, recall that, as Bonacich \citeyearpar[157]{bonacich1991simultaneous} notes, we can solve for $C^B_g$ in~\ref{eq:bon2} and $C^B_p$ in~\ref{eq:bon1} and substitute the respective solutions into \ref{eq:bon1} and~\ref{eq:bon2}. Matrix-algebraic reduction of the resulting equations would show that the Bonacich dual centralities can also be obtained as solutions to the eigensystem:

@@

Note the formal similarity (and key differences) between these equations and the Bonacich two-mode centralities in equations~\ref{eq:bon3} and~\ref{eq:bon4}. Both extract individual and group centralities as eigenvectors of the one-mode projection of the original affiliation matrix: $AA^T$ for people and $A^TA$ for groups. The difference is that the reflective centralities pre-multiply the affiliation matrix and its transpose by the inverse of the first-order centralities of the nodes in each mode before computing the eigenvalue decomposition, essentially normalizing the one-mode projection by the degrees of both sets of nodes.

@@

Now, as has been noted by two-mode data analysts in other contexts \citep[e.g.,][]{faust2005using} it turns out that this ``double-pre-weighting'' of each cell entry by the degrees of each mode (e.g., the row and column sums of the original affiliation matrix) is precisely that used for the CA of a two-mode matrix.\footnote{More accurately, cells are weighted by the inverse of the square root of the product of the row, and column sums \citep[e.g.,][124]{faust2005using}.} And indeed, the limit reflective centralities obtained from equations~\ref{eq:dam1} and~\ref{eq:dam2}---which will be given by the eigenvectors corresponding to the \textit{second} largest eigenvalue of the resulting solution---will be equivalent to the row and column scores corresponding to the first non-trivial dimension (for people and groups respectively) obtained from a simple CA of the original affiliation matrix \citep[398, eq. 9.17]{fouss2016algorithms}.\footnote{As \citet[126]{faust2005using} notes, the first non-trivial CA dimension is also given by the eigenvector corresponding to the second-largest eigenvalue of $D_p^{-1/2}AD_g^{-1/2}$.} In fact, the method of ``reflections'' is just a re-discovery of the older idea of reciprocal averaging \citep{hill1973reciprocal}, which is yet another way of computing row and column scores for elements in a two-mode matrix that are substantively equivalent to those obtained by (the first non-trivial dimension of) CA \citep{mealy2019interpreting}.

Degree-pre-weighting does not radically alter the nature of the data. Indeed, it corresponds to moving from sums to averages, thus ``adjusting'' for the influence of person-activity and group size---a seemingly perennial issue in two-mode data analysis \citep[159ff]{bonacich1991simultaneous}. This can be seen by the fact that Equations~\ref{eq:mealy1} and~\ref{eq:mealy2} show that, substantively, what the degree pre-weighting does is that, for people, co-memberships count for more in determining interpersonal similarity when the group in question is small than when it is large. On the group side, shared members who do not have many affiliations count more in determining intergroup similarity than those with many affiliations.

Accordingly, it would make no sense to call the Bonacich scores $C^B_p$ and $C^B_g$ from equations~\ref{eq:bon3} and~\ref{eq:bon4} ``centrality scores'' but fail to hold that designation from $C^R_p$ and $C^R_g$, given that the mathematics are not only ``similar'' \citep[162]{bonacich1991simultaneous} but formally identical. In the Bonacich case, an eigendecomposition of the one-mode projection of the two-mode network; in the reflective case, an eigendecomposition of the degree-weighted one-mode projection of the same network. Accordingly, the CA of a two-mode network will return---along the first dimension---a rank-ordered score for nodes in each mode that meets all the conditions for qualifying as a centrality measure for two-mode networks. Particularly, the CA centrality retains the duality property \citep[128]{faust2005using}, as the (average) centrality of people is a function of the (average) centrality of the groups they belong to. The (average) centrality of groups, in turn, is a function of their members' (average) centrality. The person and group score ranks given by the first CA dimension are thus better thought of as \textit{activity and group-size normalized} versions of the familiar Bonacich \citeyearpar{bonacich1991simultaneous} dual centralities for two-mode networks.

@@

\section{Illustrative Analysis of the Southern Women Data} \label{sec:anal}
\subsection{Row and Column Re-ordered Affiliation Matrices}
Figures~\ref{fig:ca-v-bon}(a) and~\ref{fig:ca-v-bon}(b) illustrate the key differences between the two versions of dual centrality $C^R$ and $C^B$. Each panel shows the original Southern Women affiliation matrix but with rows and columns re-ordered according to the first CA dimension in (a) and the \citet{bonacich1991simultaneous} centrality in (b). In the plot, a cell entry is colored red if it has a one in the original affiliation matrix and is white when the corresponding entry is zero. As we can see, the two row-column-reshuffled matrices have appreciably distinct structures, with the $C^R$ re-ordered affiliation matrix having a block-diagonal structure and the $C^B$ re-ordered affiliation matrix having a triangular structure. Accordingly, the $C^R$ centrality rankings reveal a dual community partition between groups of women who selectively attend two groups of events (on the top-right and bottom-left of the plot). The traditional eigenvector centrality re-ordering, on the other hand, reveals a classic core-periphery partitioning \citep{borgatti2000models}, with a group of highly active women (on the top-left) who selectively co-participate in highly attended events (on the top right-hand side of the plot) and less active women (in the bottom half) who go to less well-attended events (on the left side). The two centralities extract qualitatively different information from the two-mode network, with $C^R$ geared toward community partitioning and $C^B$ more focused on finding a core of well-connected people and groups.
